A Switching Planner for Combined Task and Observation Planning

### 2010-11-25 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
the POP stuff seems now like it is a detail. I don't see what he compilation is, there is an optimisation problem there. Is there something we can cite for that?

We now describe our continual planning approach that switches between
sequential and decision-theoretic planning sessions.

Sequential planning proceeds in a continual manner until the
verification of an assumption about runtime state is scheduled. At
that point the system switches to a decision-theoretic planner that is
used to tailor sensory processing by the robot to test the validity of
the assumptions made so far.

In this work, the problem at hand is a POMDP described in DTPDDL. From
hereon, we restrict our attention to POMDPs with deterministic
actions. Therefore it is worthwhile noting that any POMDP with
stochastic actions can be compiled into an equivalent
deterministic-action POMDP, where all the original action uncertainty
is expressed in the starting-state
distribution~\cite{ng:Jordan:2000}. In our setting, that of
finite-horizon planning, such a compilation yields a finite flat
representation of the original POMDP, only with deterministic
actions.

@@

In a decision-theoretic session, the switching planner solves a sensing
problem in an abstract processes defined in terms of the assumptive
actions in the current sequential plan, and the $(\pp{:init})$ declaration
from the original problem description.

The focus for decision-theoretic planning is on sensing, rather than
the achievement of utility according to the given objectives. We
therefor give the process in this session a different reward model, which
reflects the value of performing sensing actions in the context of the
trace proposed for the preceding sequential deliberation. In detail,
all positive rewards from the initial POMDP are removed. Then, for
each $\assumptiveS{i}$ action scheduled in the current sequential
plan, we have a {\em dual} $\assumptiveDT{i}$ so that
$\forall\state\in\states$:

@@

\noindent Where $\$(T_i)$ is a positive numeric quantity which
captures the reward the agent receives for correctly jettisoning a
false assumption. The added dual actions are propositionally trivial
-- i.e., $\forall\state\in\states$ $\poss(\assumptiveDT{i})
\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  Also, for the action,
$\action_i$, whose scheduled execution triggered the switch to the
decision-theoretic session, we include another propositionally trivial
action $\actions.\poss(\action_i)$ with the reward property:

@@

\noindent At this point, decision-theoretic planning would proceed in an
abstraction of the environment that does not contain a $\pp{cup}$,
where the inclusion (exclusion) of POMDP actions is the same as we
defined it for the sequential session.

In a second step, we iteratively refine the above{\em relaxed}
declaration by adding terms from the original statement of
$(\pp{:init})$ while the resulting state space remains of a
practicable size. Of course, it is expensive to measure size exactly,
and in practice we estimate this to be $2^{|\propositions^\#|}$, where
$\propositions^\#$ is the set of propositions in the abstraction that
were not assigned the {\em unknown} value \#. Terms are proposed for
addition because they are probabilistic, and because they were omitted
in the {\em relaxed} visitation characterised by {\em assumptive}
actions in the sequential plan. In each iteration, we add one
candidate term with the largest number of parent terms already in the
candidate tree.

>>>>

We describe our continual planning approach that switches between
sequential and contingent sessions. Continual planning and execution
proceeds in a deterministic-action POMDP described in DTPDDL. During a
sequential session, planning is performed by a ``classical''
system,\footnote{That is, a system designed to solve fully observable
deterministic tasks.}  and execution proceeds according to the {\em
trace} computed by that system. Taking the form of a classical plan,
the {\em trace} specifies a sequence of POMDP actions that achieve the
agent's objectives in a deterministic approximation, i.e., {\em
determinisation}, of the problem at hand. More precisely, the trace is
an interleaved sequence of POMDP actions and {\em assumptive}
actions. The latter correspond to assumptions the planner makes about
the truth value of {\em runtime state variables} -- e.g. that a box of
cornflakes is located in the kitchen at the third plan step. Another
kind of assumption, called an {\em applicability} assumption, is made
when the trace includes an action $\action$ that is not applicable
with probability $1$ at the belief-state
\bstate\ that the system is projected to be in when \action\ is executed, 
i.e., $\exists\state\in\states\;
\bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
scheduling
\action,  the serial planner makes an assumption about the
observability of the precondition $\poss(\action)$.

Our approach always begins with a sequential session. Non-assumptive
actions from the trace are executed until the applicability of the
next scheduled action, here \action, is questionable. A contingent
session then begins, which tailors sensory processing by the agent to
perceive whether the assumptions made in the trace hold. For each
active assumption in the trace, we add a {\em disconfirm} action to
the POMDP whose execution is rewarding (resp. costly) if the
assumption is false (resp. true). We also add one {\em confirm} action
that is rewarding (resp. costly) to execute if $\poss(\action)$ is
true (resp. false). If execution of the contingent plan applies a
disconfirm action, then a new sequential session begins at the current
underlying belief-state. If the confirm action is executed, the
sequential session is resumed, and \action\ is executed. Otherwise,
control rests in the continual session.
%%
Finally, because contingent planning is only practical in relatively
small POMDPs, contingent sessions proceed in an abstract decision
process determined by the current trace. That abstract process is
constructed by first omitting all variables from the model that are
not featured in the trace, and by then iteratively refining that
abstraction while the resulting model is of a practicable size.

@@

In a contingent session our switching planner solves a sensing problem
in an abstract processes defined in terms of the assumptive actions in
the current sequential plan, and the $(\pp{:init})$ declaration from
the original problem description.

The focus for contingent planning is on sensing, we therefore give the
process in this session an augmented reward model.  This reflects the
value of performing sensing actions in the context of the trace
proposed by the preceding sequential deliberation. First, all rewards
from the original POMDP are retained. Then, for each $\assumptiveS{i}$
action scheduled by the current trace, we have a {\em dual}
$\assumptiveDT{i}$ so that $\forall\state\in\states$:

@@

\noindent Where $\$(T_i)$ is a positive numeric quantity which
captures the reward the agent receives for correctly rejecting a
false assumption. The added dual actions are propositionally trivial
-- i.e., $\forall\state\in\states$ $\poss(\assumptiveDT{i})
\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  
%%
In order to reduce the scope of possible rejections, we compile the
serial trace into a partially ordered plan (POP). If $\action_i$ is
the sequentially scheduled execution that triggered the switch to the
contingent session, then we discard assumptive duals that are not
ordered before $\action_i$ in that POP.
%%
Also for the action $\action_i$, we include another propositionally
trivial action $\actions.\poss(\action_i)$ with the reward property:

@@

\noindent At this point, decision-theoretic planning would proceed in an
abstraction of the environment that does not contain a $\pp{cup}$,
where the applicability of POMDP actions is the same as we
defined it for the sequential session.

In a second step, we iteratively refine the above{\em relaxed}
declaration by adding terms from the original statement of
$(\pp{:init})$ while the resulting state space remains of a
practicable size. Of course, it is expensive to measure size exactly,
and in practise we estimate this to be $2^{|\propositions^\#|}$, where
$\propositions^\#$ is the set of propositions in the abstraction that
were not assigned the {\em unknown} value \#. Terms are proposed for
addition because they are probabilistic, and because they were omitted
in the {\em relaxed} visitation characterised by {\em assumptive}
actions in the sequential plan. In each iteration, we add one
candidate term with the largest number of parent terms already in the
candidate tree.
